\chapter{Introduction}

Games like Chess, Shogi, Go, and Checkers are games in which each player has perfect information regarding the state of the game. According to \cite{MYCIELSKI199241}, perfect information means that there is no hidden information between the players, each time only one of the player moves, that the game depends only on their choices, every player remembers the past, and in principle they know all the possible futures of the game.

Throughout history, these games have been benchmarks for intelligence. That is why numerous attempts have been made to beat humans in these games. DeepBlue is the first artificial intelligence to defeat a grandmaster in the game of chess in a professional setting \cite{hsu2002behind}.

\section{Background of the Study}

Artificial intelligence (AI) has made significant strides in recent years, achieving groundbreaking results in areas previously thought to require human intuition and expertise. Notable among these advancements is AlphaZero, an AI developed by DeepMind that has demonstrated an ability to master complex games such as Chess, Go, and Shogi \cite{silver2017masteringchessshogiselfplay}. By employing deep reinforcement learning and self-play, AlphaZero has set new standards in artificial intelligence by learning game strategies autonomously without any human input beyond the basicwrules of each game. This breakthrough has inspired researchers to explore applications of AlphaZero’s learning model in diverse domains.

One such potential application is in the game of Damath. Damath is a Filipino educational board gamewthat combines the principles of mathematics with the strategic nature of a board game. The game is similar to Checkers but integrates mathematical operations into each move, requiring players to use mathematical strategies along with spatial and strategic thinking. As a tool for enhancing students' mathematical and analytical skills, Damath has become popular in Philippine schools. 

Applying AlphaZero to Damath presents an opportunity not only to advance AI applications in local games but also to explore new approaches to educational AI, contributing to both the fields of AI research and educational technology.

The original AlphaZero paper makes use of Convolutional Neural Networks  particularly the ResNet architecture to develop a competitive model for games such as Chess, Shogi, and Go. However, recent advances in Computer Vision, have shown that Vision Transformers can beat Convolutional Neural Networks (\cite{dosovitskiy2021imageworth16x16words}), which inspired the researchers to determine the feasibility of this architecture and see if they can train such model to become competitive in the game of Damath.

\section{Problem Statement}

To the extent of the researchers' knowledge, most programs that play Damath rely on programmed rules or decision trees that do not fully explore the strategic potential of the game. This limitation results in AI opponents that are relatively easy to predict and exploit, reducing the educational value and engagement level for players. An AlphaZero-based model could provide a self-learning AI capable of adapting to the unique strategies and mathematical complexities of Damath, resulting in a more challenging and dynamic opponent.

This research explores the feasibility of training a model with a Vision Transformer architecture using the AlphaZero framework, and determine if it can learn various strategies in the game of Damath without prior human knowledge beyond the rules of the game. 

\section{Objectives of the Study}

The primary objective of this study is to develop and evaluate an AlphaZero-based AI model tailored to the game of Damath, more specifically integer Damath. Specific objectives include:
\begin{enumerate}
    \item  To train a model that is proficient in the game of Damath, without prior human knowledge.
    \item To evaluate the performance of the AlphaZero-based Damath AI by testing it against human players and against itself.
    \item To analyze the AI’s learning process and behavior, examining whether it can effectively learn and adapt to optimal Damath strategies without external guidance.
    % \item To assess the educational impact of the AlphaZero-based Damath AI on players, particularly in terms of enhancing their mathematical and strategic thinking abilities.
\end{enumerate}

\section{Significance of the Study}

The significance of this study lies in its potential contributions to both artificial intelligence research and educational gaming. From a technical perspective, adapting AlphaZero to a game like Damath requires innovations in handling mathematical operations as part of the AI’s decision-making process, a feature not present in typical board games like Chess or Go. Success in this endeavor could provide insights into the adaptability of AlphaZero in various educational and non-standard game environments, offering a model for future AI research in educational games.

From an educational perspective, this study could improve the effectiveness of Damath as a teaching tool, offering students a more challenging and interactive experience. An advanced AI opponent could encourage students to think critically, refine their strategies, and improve their mathematical skills, contributing to a deeper understanding and appreciation of mathematics through play.

\section{Scope and Limitations}

This study will focus on adapting the AlphaZero framework specifically to the rules and gameplay of Damath. There are many variations to the game of Damath, this study will only focus on integer Damath. Additionally, given computational constraints, the model's training process may be limited to a defined number of games or iterations. Limitations of the study include computational resources, as AlphaZero requires significant processing power for optimal training results.

\section{Organization of the Study}

The succeeding chapters of this study are organized as follows:
\begin{itemize}
    \item Chapter 2: Preliminaries – Discusses previous works and studies on AlphaZero and Damath.
    \item Chapter 3: Methodology – Details the approach, model design, and implementation process for developing the AlphaZero-based Damath AI.
    \item Chapter 4: Results and Discussion – Presents and analyzes the results of the AI’s performance.
    \item Chapter 5: Conclusion and Recommendations – Summarizes the study’s findings and suggests areas for future research.
\end{itemize}

In the following chapters, we delve deeper into the research questions and review the current state of AI applications in games, setting the foundation for this investigation into AlphaZero’s applicability to Damath.