\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgement}

The authors thank Professor Lee Javellana for his patience, guidance, and support throughout this study.  We also thank Mr. Basanes for voluntarily imparting his knowledge in the game of Damath. This study would not be possible without his expertise.

In addition, we would like to thank our family. The authors thank Mr. Bernard and Mrs. Merryl Meneses for their constant support and for leading us to communicate with Mr. Basanes. We also thank Mr. David and Mrs. Virna Bador for their constant and everlasting support throughout our time in college.

Special thanks are due to our friends who made our undergraduate years bearable and fun, Joana Mariel Santiago, Hannah Kathrina Ramos, Charrie Anne Nacor, Lester Ignacio, Andre Bryant Bagalso, Grandemir Baysa-Pee and Zed De Guzman.

% We also thank Alpha Phi Omega - Gamma Pi Chapter for 

% We also thank Common Ground and their staff for being a good host throughout our studies. We would not be able to work comfortably in our study without the warm ambiance and ample amount of coffee they provide to us every night and day.

Lastly, we also thank the University of the Philippines-Baguio, for molding us to what we are today. As scholars of the Philippines, the university has shaped us to be as sharp and critical as we are now.


% I would like to thank the members of my thesis committee for their help in preparation of this work – Niles Caulder, without whom I would have been doomed to never complete it, Kimiyo Hoshi, who helped to shed new light on many of my ideas, Pamela Isley, with whom I often disagree but who inspires me to be better, Raymond Palmer, who had no small part to play in the formation of the idea, and Kent Nelson, who always
% had golden advice.

% Special thanks are due to the friends and colleagues who made this work possible. Jimmy Olsen and Pete Ross were invaluable both as friends and as sounding boards for some of my more outlandish ideas. Jack Knight, who I met only briefly, was a major influence, and I’m glad we were able to help each other.

% The author gratefully acknowledges the support for this work offered by S.T.A.R. Laboratories under grant award number 3X29YZ4A, and by the Theodore S. Kord Fellowship. Any views and conclusions contained herein are those of the author, and do not necessarily represent the official positions, express or implied, of the funders.